---
title: "Homework 3 template"
author: "Ryan Oldridge"
format: gfm
editor: visual
---

# Data input/output

loading tidyverse and reading the data file and making it an object

```{r, message=FALSE}
library(tidyverse)

fake_data <- read_csv(file = "fake_date.csv")
```

```{r}
read_csv("fake_date.csv")
```

```{r}
print("fake_data")
```

# Investigating the properties of data frames

Use two different functions that both give some kind of summary or overview of the data frame. Which one do you think is more useful?

```{r}
str(fake_data)
```

```{r}
glimpse(fake_data)
```

glimpse is more useful here because it shows the data clearly laid out in an organized way.

# Manipulating data frames

Create a new column of your data frame by performing some kind of arithmetic operation on an existing column.

```{r}
fake_data$newsomething <- fake_data$something + 2 

glimpse(fake_data)
```

created a new column by adding 2 to the column "something"

# Working on columns

Calculate the mean of a numeric column in your data frame.

```{r}
mean(fake_data$newsomething)
```

found that the mean of the new column i made is 7.5

# Accessing elements of data frames

```{r}
cats <- data.frame(coat = c("calico", "black", "tabby", "calico"),
                   weight = c(2.1, 5.0, 3.2, 4.4),
                   likes.string = c(TRUE, FALSE, TRUE, TRUE))
```

```{r}
cats[1]
```

shows the first list which is the coat of the cats

```{r}
cats[[1]]

```

listed the first column as a vector

```{r}
cats$coat
```

listed the values within the first column which happen to be the coats.

```{r}
cats[1, 1]
```

this called the first value of the first column

```{r}
cats[ , 1]
```

this gave me every value of the first column

```{r}
cats[1, ]
```

gives me all of the first row of each column
